---
title: "Cassowary case study"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cassowary case study}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
pacman::p_load(tidyverse, gt)
theme_set(theme_bw())
library(monitoR)
```

# Load data
```{r}
data(cassowary)
cassowary |> glimpse()
```


# Adding the grid
We add a 4x4 grid - decide to make a bit nicer than the default grid. Note that there are two zones denotes left and right (@fig-grid).

```{r}
grid <- create_grid(
  range(cassowary$X),
  range(cassowary$Y),
  dim = c(4, 4)
)
grid <- grid |> 
  mutate(
    left = case_when(
      left == 74 ~ 0, 
      left == 195.25 ~ 150, 
      left == 316.5 ~ 300, 
      left == 437.75 ~ 450, 
      TRUE ~ left
    ),
    right = case_when(
      right == 74 ~ 0, 
      right == 195.25 ~ 150, 
      right == 316.5 ~ 300, 
      right == 437.75 ~ 450, 
      right == 559 ~ 600, 
      TRUE ~ right
    )
  )
grid$zone <- "right"
grid$zone[c(1,2,5,6,9,10,13,14)] <- "left"
cassowary_grid <- cassowary |> add_grid(grid)
```

```{r}
#| fig-cap: Grid plot the cassowary dataset. 
#| label: fig-grid
plot_grid(grid, cassowary_grid, zone_fill = TRUE) + 
  theme(legend.position = "none")
```

# Cassowary side
We have two cassowaries - Martina and Jeffrey. They are let out into on of the sides. We have information about which side they are in. 

```{r}
martina <- cassowary_grid |> 
  filter(cassowary == "Martina")
jeffrey <- cassowary_grid |> 
  filter(cassowary == "Jeffrey")
jeffrey
```

@fig-martina gives the positions for Martina and also the side that she is recorded on, while @fig-jeffrey gives the equivalent for Jeffrey. 

::: {.callout-note}
Need to discuss with EF. 
:::

```{r}
#| fig-cap: Martina data
#| label: fig-martina
grid |>
  ggplot2::ggplot() +
  ggplot2::geom_rect(
    ggplot2::aes(xmin = left, xmax = right, ymin = bottom, ymax = top),
    col = "black", alpha = 0.1, fill = NA
  ) +
  ggplot2::geom_point(
    ggplot2::aes(X, Y, col = martina_side),
    data = martina
  ) + 
    harrypotter::scale_colour_hp_d("Ravenclaw")
```

```{r}
#| fig-cap: Jeffrey data
#| label: fig-jeffrey
grid |>
  ggplot2::ggplot() +
  ggplot2::geom_rect(
    ggplot2::aes(xmin = left, xmax = right, ymin = bottom, ymax = top),
    col = "black", alpha = 0.1, fill = NA
  ) +
  ggplot2::geom_point(
    ggplot2::aes(X, Y, col = jeffery_side),
    data = jeffrey
  ) + 
    harrypotter::scale_colour_hp_d("Ravenclaw")
```


@fig-prop-left gives the proportion of time that each cassowary is in each grid for the left side. 

```{r}
#| fig-cap: Proportion of time in each grid
#| label: fig-prop-left
cassowary_grid |> 
  filter(zone == "left") |> 
  count(grid, cassowary) |> 
  mutate(
    N = sum(n), .by = cassowary, 
  ) |> mutate(
    p = n / N, 
    grid = factor(
      grid, 
      levels = c(13, 14, 15, 16, 9, 10, 11, 12, 
    5, 6, 7, 8, 1, 2, 3, 4)
    )
  ) |> 
    ggplot(aes(cassowary, p, fill = cassowary)) + 
    geom_col() + 
    facet_wrap(~grid, ncol = 2) + 
    harrypotter::scale_fill_hp_d("Ravenclaw")
```

```{r}
#| fig-cap: Proportion of time in each grid
#| label: fig-prop-right
cassowary_grid |> 
  filter(zone == "right") |> 
  count(grid, cassowary) |> 
  mutate(
    N = sum(n), .by = cassowary, 
  ) |> mutate(
    p = n / N, 
    grid = factor(
      grid, 
      levels = c(13, 14, 15, 16, 9, 10, 11, 12, 
    5, 6, 7, 8, 1, 2, 3, 4)
    )
  ) |> 
    ggplot(aes(cassowary, p, fill = cassowary)) + 
    geom_col() + 
    facet_wrap(~grid, ncol = 2) + 
    harrypotter::scale_fill_hp_d("Ravenclaw")
```

# Diversity measures

## Entropy
We see that Martina utilises more of the grid than Jeffrey (@tbl-ent)
```{r}
#| tbl-cap: Entropy for each cassowary
#| label: tbl-ent
cassowary_grid |> 
  group_by(cassowary) |> 
  summarise(entropy = calc_entropy(grid)) |> 
  gt() |> 
  fmt_number(decimals = 4)
```

@fig-b-ent gives the entropy for each behaviour for each cassowary. 

```{r}
#| fig-cap: Behaviour entropy
#| label: fig-b-ent
cassowary_grid |> 
  group_by(cassowary, behaviour) |> 
  summarise(entropy = calc_entropy(grid)) |> 
  filter(entropy > 0) |> 
  group_by(behaviour) |> 
  mutate(m = max(entropy)) |> 
  ungroup() |> 
  mutate(
    behaviour = fct_reorder(behaviour, m)
  ) |> 
  ggplot(aes(entropy, behaviour, fill = cassowary)) + 
  geom_col(position = "dodge") + 
  harrypotter::scale_fill_hp_d("Ravenclaw") + 
  labs(x = "Entropy", y = NULL)
```


```{r}
get_zone_object(grid, martina) |> calc_ei()
get_zone_object(grid, jeffrey) |> calc_ei()
get_zone_object(grid, martina) |> calc_spi()
get_zone_object(grid, jeffrey) |> calc_spi()
```